---
title: "covid 19"
author: "Daniella Parra"
date: "Summer 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

We can see from the first table how many people we start off. In the females we have 1,831 and 25 dead. 

In the males we have 1,411 and 44 dead. 

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

#### Plot barplots tigerstats way

```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")

```

In the barchart above we can see that blue is female and light purple is males, and in the female on the left we can see that they are roughly around 58% alive, and on the right they are roughly around 35% dead. 

On the male side we can see that they are around 42% less alive compared to the females, we also have to consider that they did start off with a little over 400 people less than the females. Although on the male side they did have more people die which makes their percentage a little bit over 60%.  


#### Plot barplots ggplot way

```{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

We can see this chart is the same as the one above, although this one is more clear and we can see what is exact. 

The bars show what we had predicted before, where the females are higher than the males on being alive, and males are higher than females on how many died. 




  
## Numerical results

```{r}
table1 <- xtabs(~sex + result, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The two results are shown above with a percentage and one for the full numbers out of 100.

For the percentage above we can see that only 1.4% have died to covid-19  in the female side leaving us with only 98.7% females alive. Going into more detail we can see out of 100, 56 females are alive and only 36 have died. 

In the males percentage we can see that only 3.0% have died which makes the bar charts correct, and leaves that with 97.0% males that are alive. Going into more detail for the males out of 100 we can see that 64 males have died and only 44 are alive which is a huge difference. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Since the p-value is 0.0012 since the p-value is so small we can reject this hypothesis because it is more unlikely to occur, due to us using sex as something that affects this virus. We have to reject this because the null hypothesis that death is independent of sex is rejected in favor of the alternative, that sex plays a role in death.

## Conclusions

We first started off by using both barcharts to determine between percentages, and out of 100. We can determine that the males had the most impact during this because they had more deaths than the females, and the females had the less deaths. 

In the end we had to reject this hypothesis and have it as an alternative because we used sex which doesn't play an actual role for death. 
